Stack Safety is a Security Property

### 2021-04-14 tex/main.tex
pass through 4

We choose to model observations explicitly, rather than simply treating
a subset of registers or memory as being observable.  This choice gives
us the flexibility to treat programs as observationally equivalent even when
their internal states may differ.
%The choice of modelling observations is an interesting one. A first approach
%would be to consider any change to the registers or memory as observable.
%However, as some parts of the stack will be considered ``private''
%this choice won't work. A relaxation of this to consider all changes
%to ``non-private'' data as observable could work in principle, but
%does not allow for an interesting weaker-but-more-efficient enforcement
%mechanism. We will come back to this point in \Cref{sec:lazy}.

In the running example of \Cref{sec:running-example},
$\WORDS$ is the
set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname,\SP\}$. Observations are writes to the
special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
  not...)}\fi, which can be thought of as a memory-mapped
  output port: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

@@

This combined step function is also partial, and traces of the machine-with-policy
will be finite iff a policy fault occurs.

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
\(\mach, \context, \ann \in \powerset \ANNS \constep \context' \in \MACHS \times \CONTEXTS
\times \ANNS \rightarrow \CONTEXTS\). We also assume for any set of annotations that we have
code maps, with \(\FUNIDS\) a set of function identifiers:
\[\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]

We lift contexts into an MP trace in much the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \ann\)}
              {\(\mach, \context, \ann \constep \context'\)}
              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
                \mpcstate{\mach'}{\pol'}{\context'}\)}

@@

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.

@@

The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
creates a trace of observations corresponding to the observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
initial observation, a machine-trace, and an observation trace:

@@

In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
to relate a machine-policy-context state triple to the MPCO-trace
induced by the step function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:

@@

In particular, we wish to formulate properties around consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''

@@

\begin{center}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\callmap\} \cup \ann) \stepstocon
             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\retmap\} \cup \ann) \stepstocon
             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\passmap ~ P\} \cup \ann) \stepstocon
             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\sharemap ~ P\} \cup \ann) \stepstocon
             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
  \judgment{}
           {\(\mach,\context,\_ \stepstocon \context\)}
\end{center}

@@

Recall that our properties use a compiler-derived code-map determine
which addresses are instructions and which functions they belong to, as well
as annotating them as calls, returns, and yields, as well as sharing information.
We now extend it with an annotation for valid entry points:
\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

>>>>

Observations are only essential for describing the extensional properties
described in Section~\ref{sec:lazy}, but for simplicity, we include them
in the machine model from the beginning.
We choose to model observations explicitly, rather than simply treating
a subset of registers or memory as being observable.  This choice gives
us the flexibility to treat programs as observationally equivalent even when
their internal states may differ.
%The choice of modelling observations is an interesting one. A first approach
%would be to consider any change to the registers or memory as observable.
%However, as some parts of the stack will be considered ``private''
%this choice won't work. A relaxation of this to consider all changes
%to ``non-private'' data as observable could work in principle, but
%does not allow for an interesting weaker-but-more-efficient enforcement
%mechanism. We will come back to this point in \Cref{sec:lazy}.

\apt{This para is out of date. Do we want to say something equivalent
  about new examples (hard, because they are C-based).}
$\WORDS$ is the
set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname,\SP\}$. Observations are writes to the
special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
  not...)}\fi, which can be thought of as a memory-mapped
  output port: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

@@

\noindent
This combined step function is also partial, and traces of the machine-with-policy
will be finite iff a policy fault occurs.

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
\(\mach, \context, \anns \constep \context' \in \MACHS \times \CONTEXTS
\times \powerset(\ANNS) \rightarrow \CONTEXTS\). We also assume a set of function identifiers
\(\FUNIDS\) and the existence of a \emph{code map} \(\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\)
that relates each code address to its owning function and set of annotations.
We lift contexts into an MP trace in much the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \anns\)}
              {\(\mach, \context, \anns \constep \context'\)}
              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
                \mpcstate{\mach'}{\pol'}{\context'}\)}

@@

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\). %, and of observations.

@@

The ``observations-of'' operator, written \(\pi_\obs\), \apt{since this isn't
  actually a projection, it is slightly confusing to use \(\pi\)} takes a machine trace and
creates a trace of observations corresponding to the observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
initial observation, a machine-trace, and an observation trace:

@@

In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
to relate a machine-policy-context state triple to the MPCO-trace
induced by the step function.\apt{?? This looks indistinguishable
  from the \(\hookrightarrow\) we've already defined.}

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:

@@

In particular, we wish to formulate properties about consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''

@@

\begin{center}
  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\callmap\} \cup \anns) \stepstocon
             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\retmap\} \cup \anns) \stepstocon
             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\passmap ~ P\} \cup \anns) \stepstocon
             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
  \judgment{\(\mach,\context,\anns \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\sharemap ~ P\} \cup \anns) \stepstocon
             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
  \judgment{}
           {\(\mach,\context,\_ \stepstocon \context\)}
\end{center}

@@

Recall that our properties use a compiler-derived code-map determine
which addresses are instructions and which functions they belong to, as well
as annotating them as calls, returns, and yields, as well as sharing information.
We now extend it with an annotation for valid entry points:
\[\ann \in \ANNS ::= \callmap + \retmap + \yieldmap +  \entmap + 
\sharemap(P \subseteq \MACHS \times \COMPONENTS) +
\passmap(P \subseteq \MACHS \times \COMPONENTS)\]
